---
title: "Assignment 4 - Heart rate, Respiration and interpersonal coordination"
author: "Asger L. Christiansen"
date: "August 20, 2019"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 4 - Heart rate, Respiration and interpersonal coordination

Physiological data (here heart rate [variability], and Respiration) are increasingly popular. Historically treated as pernicious noise to be regressed out of neuro-imaging data, there is now increasing research on how these signals tell us something important about cognition and beyond being just a signal of cognitive processes also impact them in interesting ways. Advanced sport science, and the quantified self movement (closely followed by marketing and communication) have hailed continuous physiological tracking as a powerful way to access and modify attitudes, habits, and performance. Further, as team coordination (in the military, in decision processes and organizational contexts) is more and more in focus, research has attempted to measure how interpersonal coordination between physiological systems might tell us something important about e.g. emotional and cognitive coordination. See references in the reading list for more on this.

In this assignment, you will learn to:
- collect physiological data
- pre-process physiological data (and grow further your mad R skills)
- model the continuous interdependence between two signals (using a multilevel model as proxy for a dynamical system approach)
- conservatively assess the presence of coordination between to signals in a controlled context

This assignment has two parts. The first part familiarizes you with heart rate, and Respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and Respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and Respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is Respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from previous years (Study1, Study2 and Study 3). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: in the first year it was self-paced joint reading; in the second year it was the tv-series conversation.

## Let's get started

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' Respiration signal and a different one of the participants' HR signal.
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3). There are also smarter packages, like cowplot and ggpubr.
- Can you eye-ball which condition if any displays more physiological coordination?

### First we read one data file and identify the procedure
- Load the file
- correctly identify all columns
- plot the data
- deal with the artifacts
- downsample the dat
- Add a column for study, group, trial and condition

```{r}
# Loading packages and data
pacman::p_load(pacman,tidyverse, stringr, stringi, groupdata2)

# Loading an example data file of the "synchronous" condition
sync_raw <- read.csv("data/Study4_G8_T2_Synchronous.csv")

## Remove outliers

### Tip, check the function below
remove_outs<- function(ts,threshold){
  sigma=sd(ts,na.rm=T)
  mu=(mean(ts,na.rm=T))
  ts[ts > (mu +
             (threshold*sigma)) ] = mu +
             (threshold*sigma)
  ts[ts < (mu -
             (threshold*sigma)) ] = mu -
    (threshold*sigma)
  return(ts)
}

threshold=2.5 # Default value at 2.5 sds from the mean


# Testing the remove_outs function
sync_2 <-sync_raw %>% 
  mutate_at(c("min","Resp1","Resp2","HR1","HR2", "ECG1", "ECG2"), remove_outs, threshold)


# We need to scale the data (different people have different hert rate ranges and the like).

# Testing whether scale() does the job

sync_3<-sync_2 %>% mutate_at(c("min","Resp1","Resp2","HR1","HR2", "ECG1", "ECG2"), scale)

## Downsample
### We create the following function to downsample



downsample <- function(df) {
  
  moose <- df %>%
  groupdata2::group(n = 1000, method = 'greedy') %>%
  dplyr::summarise(
   time = mean(TimeMs,na.rm=T),
   HR1 = mean(HR1,na.rm=T),
   HR2 = mean(HR2,na.rm=T),
   Resp1 = mean(Resp1,na.rm=T),
   Resp2 = mean(Resp2,na.rm=T))
   
  moose$rowname <- (1:nrow(moose))
  
    return(moose)
}

# Testing the function

sync_4 <- downsample(sync_3)



pacman::p_load(pacman,tidyverse, stringr, stringi)

## Plotting examples of Heart Rate data at various stages of preprocessing
raw_plot <- sync_raw$HR1 %>% matplot(type="l") # Raw data
raw_plot

no_outliers_plot <- sync_2$HR1 %>% matplot(type="l") # Outliers removed
no_outliers_plot

scaled_plot <- sync_3$HR1 %>% matplot(type="l")# Data scaled
scaled_plot

downsampled_plot <- sync_4$HR1 %>% matplot(type="l")# Downsampled data
downsampled_plot



# Now we add the group, trial, condition to the cleaned up, scaled, downsampled data

## Using the following filename as an example:

"data/Study4_G8_T1_TurnTaking.csv"

### We create the following function to do it with:

read_title <- function(filename) {
    #Parse filename to extract Study, group, trial and condition
    #Extracting study:
    study <- str_extract(filename, pattern = "\\d")
    #Splitting it once by the letter G
    split_1 <- str_split(filename, pattern = "G", simplify = T)
    #Extracting Group nr from the second string in "split_1"
    group <- str_extract(split_1[2], "\\d")
    #Splitting the filename again by the letter T
    split_2 <- str_split(filename, pattern = "T", simplify = T)
    #Extracting Trial from the second string in "split_2"
    trial <- str_extract(split_2[2], "\\d")
    #Extracting condition from the second string of split_3
    condition <- split_1[2] %>% str_replace_all(pattern = "\\d", "") %>% str_replace_all(pattern = ".csv", "") %>% str_replace_all(pattern = "_", "") %>% str_replace(pattern = "T", "")
    moose <- data.frame(study, group, trial, condition)
    
    return(moose)
}

# Testing the function:

read_title("data/Study4_G8_T1_TurnTaking.csv")


```


## Now we are ready to go to load and pre-process all files

Go through all the files (with a function passed onto map_df), check which files should be excluded, if any, and save the pre-processed time-series

A couple of tips:
- looping is oh so slow. Making a function and using Map/Map_df is your salvation.
- each study restarts the group numbering, so you should make sure to change that (e.g. 100 * Study + Group)
- you need to make sure all the data are meaningful or something has to be removed. Plotting is your friend. E.g. "Study1_G1_T1_sync_raw" has one bad Respiration signal. We could replace it with NAs


```{r}
# Reloading packages (to make working within a single chunk easier)
pacman::p_load(pacman,tidyverse, stringr, stringi, groupdata2)

# Creating a pre-process function which does all of the above
pre_process <- function(filename) {
  # Creating two objects, one using read.csv, the other using read_title
  df <- read.csv(filename)
  numbers <- read_title(filename)
  # Setting threshold for remove_outs
  threshold = 2.5
  
  # Mutating df with remove_outs and scale
  df  <- df %>%   mutate_at(c("Resp1","Resp2","HR1","HR2", "ECG1", "ECG2"), remove_outs, threshold)
  df <- df %>%    mutate_at(c("Resp1","Resp2","HR1","HR2", "ECG1", "ECG2"), scale)
  # Correcting some name issues between studies
  print(is.na(numbers$study))
  if (numbers$study != 4){df <- df %>% rename(TimeMs = time)}
  # Downsampling 1000 times
  df <- downsample(df)
  
  moose <- cbind(numbers, df)
    
    return(moose)
}

# Testing the function on a single file
example <- pre_process("data/Study4_G8_T3_Conversation.csv")

# Setting working directory (briefly) to /data, and applying pre_process() to all the files, creating one big dataframe called "phys_data"
setwd("C:/Users/Asger/Desktop/Cognitive Science BA/3. Semester/ExpMeth 3/Assignments/Assignment-4/data/"); phys_data = list.files(pattern = ".csv") %>% purrr::map_df(pre_process); setwd("C:/Users/Asger/Desktop/Cognitive Science BA/3. Semester/ExpMeth 3/Assignments/Assignment-4")

# Writing the result of this as a .csv
write.csv(phys_data, "Pre-processed Physiological Data.csv")


phys_data <- read.csv("Pre-processed Physiological Data.csv")
phys_data[,1] <- NULL

# Now we need to make sure all the data are meaningful or something has to be removed. For the purposes of this exercise, we will only be making models based on study 4. As a result, we have only plotted and examined the data from study 4


# Plotting time:
## Isolating study 4
s4 <- phys_data %>% filter(study == 4)

## looping (For heartrate) 
for(i in 1:8) {
gru <- s4 %>% filter(group == i)
plott <- ggplot(gru, aes(x = time, y = value)); plott <- plott + facet_wrap(~ condition, scales = "free_x") + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
nam <- paste("A", i, sep = "")
assign(nam, plott)
}

# Checking the plots
A1
A2
A3 # This one is weird
A4
A5
A6 # This one also has weird plots
A7 # There seems to be some flatlinging here
A8 # Outliers and weird plots

# loop-de-loop (For respiration) 
for(i in 1:8) {
gru <- s4 %>% filter(group == i)
plott <- ggplot(gru, aes(x = time, y = value)); plott <- plott + facet_wrap(~ condition, scales = "free_x") + geom_line(aes(y = Resp1, color = "Resp1"), size = 1) + geom_line(aes(y = Resp2, colour = "Resp2"))
nam <- paste("B", i, sep = "")
assign(nam, plott)
}

# Checking the plots
B1
B2
B3
B4 # Suspect data 
B5 
B6 
B7 
B8 

# List of specific files I don't like
# > Group 3 HR - Conversation, Movement, Synchronous, TurnTaking 
# > Group 6 HR - Synchronous, TurnTaking
# > Group 7 HR - Movement, MovementCoop, Conversation, TurnTaking
# > Group 8 HR - MovemntCoop
# > Group 4 Resp - Movement, Turntaking

# Lets have a look at the weird ones: 
# HR - Group 3 - Conversation
HR_3_Con <- s4 %>% filter(group == 3) %>% filter(condition == "Conversation")
Plot_HR_3_con <- ggplot(HR_3_Con, aes(x = time, y = value)); Plot_HR_3_con <- Plot_HR_3_con + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_3_con #<--- To be terminated, HR Group 3, Conv, HR1
# HR - Group 3 - Movement
HR_3_Mov <- s4 %>% filter(group == 3) %>% filter(condition == "Movement")
Plot_HR_3_mov <- ggplot(HR_3_Mov, aes(x = time, y = value)); Plot_HR_3_mov <- Plot_HR_3_mov + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_3_mov #<--- To be terminated, HR Group 3, Mov, HR1
# HR - Group 3 - Synchronous
HR_3_Syn <- s4 %>% filter(group == 3) %>% filter(condition == "Synchronous")
Plot_HR_3_syn <- ggplot(HR_3_Syn, aes(x = time, y = value)); Plot_HR_3_syn <- Plot_HR_3_syn + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_3_syn #<--- To be terminated, HR Group 3, Syn, HR1
# HR - Group 3 - TurnTaking
HR_3_Tt <- s4 %>% filter(group == 3) %>% filter(condition == "TurnTaking")
Plot_HR_3_tt <- ggplot(HR_3_Tt, aes(x = time, y = value)); Plot_HR_3_tt <- Plot_HR_3_tt + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_3_tt #<--- To be terminated, HR Group 3, TT, HR1
# HR - Group 6 - Synchronous
HR_6_Syn <- s4 %>% filter(group == 6) %>% filter(condition == "Synchronous")
Plot_HR_6_syn <- ggplot(HR_6_Syn, aes(x = time, y = value)); Plot_HR_6_syn <- Plot_HR_6_syn + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_6_syn #<--- To be terminated, HR Group 6, Syn, HR1
# HR - Group 6 - TurnTaking
HR_6_Tt <- s4 %>% filter(group == 6) %>% filter(condition == "TurnTaking")
Plot_HR_6_tt <- ggplot(HR_6_Tt, aes(x = time, y = value)); Plot_HR_6_tt <- Plot_HR_6_tt + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2")) #<--- To be terminated, HR Group 6, TT, HR1
Plot_HR_6_tt
# HR - Group 7 - Movement
HR_7_Mov <- s4 %>% filter(group == 7) %>% filter(condition == "Movement")
Plot_HR_7_mov <- ggplot(HR_7_Mov, aes(x = time, y = value)); Plot_HR_7_mov <- Plot_HR_7_mov + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_7_mov #<--- To be terminated, HR Group 7, Mov, HR2
# HR - Group 7 - MovementCoop
HR_7_MovC <- s4 %>% filter(group == 7) %>% filter(condition == "MovementCoop")
Plot_HR_7_movc <- ggplot(HR_7_MovC, aes(x = time, y = value)); Plot_HR_7_movc <- Plot_HR_7_movc + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_7_movc #<--- To be terminated, HR Group 7, MovC, HR2
# HR - Group 7 - Conversation
HR_7_Con <- s4 %>% filter(group == 7) %>% filter(condition == "Conversation")
Plot_HR_7_con <- ggplot(HR_7_Con, aes(x = time, y = value)); Plot_HR_7_con <- Plot_HR_7_con + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_7_con # We've decided to keep this, despite it not being that great. It isn't horrible either and we need data.
# HR - Group 7 - TurnTaking
HR_7_Tt <- s4 %>% filter(group == 7) %>% filter(condition == "TurnTaking")
Plot_HR_7_tt <- ggplot(HR_7_Tt, aes(x = time, y = value)); Plot_HR_7_tt <- Plot_HR_7_tt + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_7_tt #<--- To be terminated, HR Group 7, TT, HR2
# HR - Group 7 - Synchronous
HR_7_Syn <- s4 %>% filter(group == 7) %>% filter(condition == "Synchronous")
Plot_HR_7_syn <- ggplot(HR_7_Syn, aes(x = time, y = value)); Plot_HR_7_syn <- Plot_HR_7_syn + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_7_syn #<--- To be terminated, HR Group 7, Syn, HR2
# HR - Group 8 - MovementCoop
HR_8_MovC <- s4 %>% filter(group == 8) %>% filter(condition == "MovementCoop")
Plot_HR_8_movc <- ggplot(HR_8_MovC, aes(x = time, y = value)); Plot_HR_8_movc <- Plot_HR_8_movc + geom_line(aes(y = HR1, color = "HR1"), size = 1) + geom_line(aes(y = HR2, colour = "HR2"))
Plot_HR_8_movc #<--- To be terminated, HR Group 8, MovC, HR2

# Resp - Group 4 - Movement
Resp_4_Mov <- s4 %>% filter(group == 4) %>% filter(condition == "Movement")
Plot_Resp_4_mov <- ggplot(Resp_4_Mov, aes(x = time, y = value)); Plot_Resp_4_mov <- Plot_Resp_4_mov + geom_line(aes(y = Resp1, color = "Resp1"), size = 1) + geom_line(aes(y = Resp2, colour = "Resp2"))
Plot_Resp_4_mov # We've decided to keep this, despite it not being that great. It isn't horrible either and we need data.
# Resp - Group 4 - TurnTaking
Resp_4_Tt <- s4 %>% filter(group == 4) %>% filter(condition == "TurnTaking")
Plot_Resp_4_tt <- ggplot(Resp_4_Tt, aes(x = time, y = value)); Plot_Resp_4_tt <- Plot_Resp_4_tt + geom_line(aes(y = Resp1, color = "Resp1"), size = 1) + geom_line(aes(y = Resp2, colour = "Resp2"))
Plot_Resp_4_tt #<--- To be terminated, Resp Group 4, TT, Resp1
# Resp - Group 4 - MovementCoop 
Resp_4_MovC <- s4 %>% filter(group == 4) %>% filter(condition == "MovementCoop")
Plot_Resp_4_movc <- ggplot(Resp_4_MovC, aes(x = time, y = value)); Plot_Resp_4_movc <- Plot_Resp_4_movc + geom_line(aes(y = Resp1, color = "Resp1"), size = 1) + geom_line(aes(y = Resp2, colour = "Resp2"))
Plot_Resp_4_movc # We've decided to keep this, despite it not being that great. It isn't horrible either and we need data. 

# Remove bad data
## We have elected to remove the following signals:

### HR Group 3, Conv, HR1
### HR Group 3, Mov, HR1
### HR Group 3, Sync, HR1
### HR Group 3, TurnTaking, HR1
### HR Group 6, Syn, HR1
### HR Group 6, TurnTaking, HR1
### HR Group 7, Mov, HR2
### HR Group 7, MovC, HR2
### HR Group 7, TurnTaking, HR2
### HR Group 7, Syn, HR2
### HR Group 8, MovC, HR2
### Resp Group 4, TT Resp1

## Replacing those signals with NA's

phys_data$HR1[phys_data$study == 4 & phys_data$group == 3 & phys_data$condition == "Conversation"] <- NA
phys_data$HR1[phys_data$study == 4 & phys_data$group == 3 & phys_data$condition == "Movement"] <- NA
phys_data$HR1[phys_data$study == 4 & phys_data$group == 3 & phys_data$condition == "Synchronous"] <- NA
phys_data$HR1[phys_data$study == 4 & phys_data$group == 3 & phys_data$condition == "TurnTaking"] <- NA
phys_data$HR1[phys_data$study == 4 & phys_data$group == 6 & phys_data$condition == "Synchronous"] <- NA
phys_data$HR1[phys_data$study == 4 & phys_data$group == 6 & phys_data$condition == "TurnTaking"] <- NA
phys_data$HR2[phys_data$study == 4 & phys_data$group == 7 & phys_data$condition == "Movement"] <- NA
phys_data$HR2[phys_data$study == 4 & phys_data$group == 7 & phys_data$condition == "MovementCoop"] <- NA
phys_data$HR2[phys_data$study == 4 & phys_data$group == 7 & phys_data$condition == "TurnTaking"] <- NA
phys_data$HR2[phys_data$study == 4 & phys_data$group == 7 & phys_data$condition == "Synchronous"] <- NA
phys_data$HR2[phys_data$study == 4 & phys_data$group == 8 & phys_data$condition == "MovementCoop"] <- NA
phys_data$Resp1[phys_data$study == 4 & phys_data$group == 4 & phys_data$condition == "TurnTaking"] <- NA

good_data <- phys_data # Obviously there may still be bad data in the other studies. We found removing that to be beyond the scope of this assignment.

# Save the data
write.csv(good_data, "Bad Data Removed.csv")

```

## Now we need to run some analysis

Let's start with a multilevel model that accounts for 
- stability (how each signal is autocorrelated)
- interpersonal dependence (each signal is dependent from the previous state of the other signal)

The data needs to be further prepared, so we can analyze both participants in the same model.
We need to turn the data into a long format:
- a column indicating own hr and one own Respiration
- a column indicating other hr and one other Respiration
- a column indicating change in hr from previous round and one in Respiration

We can then run an analysis where change is a function of one's previous state (stability, see slides), and the other's previous state (coupling). Make sure to:
- set up the most interesting contrasts: how do these parameters vary by condition? which condition should be baseline?
- set up the right random effects.
- N.B. the model will be slow. Make sure it works on a subset of the data first!

Bonus question: what if we include an additional layer? Is my heart rate just adjusting to yours, or also to how much you are adjusting to mine?
- to start answering this we can add a column indicating the previous change in hr in the other and one in Respiration
- we can then build on the previous models by also adding the previous change in the other


```{r}
# Generate a column for each: previous HR1, HR2, Resp1, Resp2
# Generate a column for each: change in HR1, HR2, Resp1, Resp2


good_data <- read.csv("Bad Data Removed.csv")

good_data <- good_data %>% 
  group_by(group, condition) %>% 
  mutate(HR1fut = lead(HR1,1), #lead does the same as lag
         HR2fut = lead(HR2,1), 
         Resp1fut = lead(Resp1,1), 
         Resp2fut = lead(Resp2,1),
         HR1change = HR1fut - HR1,
         HR2change = HR2fut - HR2,
         Resp1change = Resp1fut - Resp1,
         Resp2change = Resp2fut - Resp2)

# Make the data long, so we can analyze both participants at the same time 


# creating long data for HR change
d_HR_change_self <- gather(good_data, participant, HR_change_self, # new vars
HR1change, HR2change) %>% #old vars
  select( # drop irrelevant vars
  time, HR_change_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#creating long data for Respiration
d_Resp_change_self <- gather(good_data, participant, Resp_change_self, # new vars
Resp1change, Resp2change) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_change_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#heart rate self
d_HR_self <- gather(good_data, participant, HR_self, # new vars
HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
  time, HR_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))


#future heart rate
d_HR_fut <- gather(good_data, participant, HR_fut, # new vars
HR1fut, HR2fut) %>% #old vars
  select( # drop irrelevant vars
  time, HR_fut, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))



#Respiration self
d_Resp_self <- gather(good_data, participant, Resp_self, # new vars
Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#future Respiration rate
d_Resp_fut <- gather(good_data, participant, Resp_fut, # new vars
Resp1fut, Resp2fut) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_fut, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))


# Making a new dataframe in which HR1/2 and Resp1/2 are switched, to generate the "other" variables
good_data_2 = good_data
good_data_2$HR1 = good_data$HR2
good_data_2$HR2 = good_data$HR1
good_data_2$Resp1 = good_data$Resp2
good_data_2$Resp2 = good_data$Resp1

#HR other
d_HR_other <- gather(good_data_2, participant, HR_other, # new vars
HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
  time, HR_other, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#Resp other
d_Resp_other <- gather(good_data_2, participant, Resp_other, # new vars
Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_other, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))


#merging dataframes
d_all = cbind(d_HR_change_self, d_HR_self, d_HR_fut, d_HR_other, d_Resp_change_self, d_Resp_self, d_Resp_fut, d_Resp_other)
phys_long = d_all %>% select(time, HR_change_self, HR_self, HR_fut, HR_other, participant, study, group, condition, trial, Resp_change_self, Resp_self, Resp_fut, Resp_other)

#saving long data
write.csv(phys_long, "Bad Data Removed, Long.csv")

```




```{r}
# Modelling time! Let's set up the data

## Reloading packages for convenience
pacman::p_load(pacman,tidyverse, stringr, stringi, lmerTest)

phys_long <- read.csv("Bad Data Removed, Long.csv")
phys_long[,1] <- NULL

## We will only analyze data from Study 4, because that is easier.
phys_4 <- phys_long %>% filter(study == 4)

## Fixed Effects

### HR_self = Stability ## Random effect: Well-trained people have different stability since HR variability is actually healthy.
### HR_other = coupling. ## Random effect: Some pairs might actually hate eachother.
### condition = condition ## Random effect: If you hate the other person, the condition doesn't matter.


## Syntax Shenanigans

### lm() implicitly adds a "1+" to the syntax, which means "Pick an intercept"
### 0+ tells lme4() to not make several intercepts. In this case, we should say "0+condition + (HR_s + HR_o):condition"
### Then take the above and do: (0+condition + (HR_s+HR_o):condition):Type

## Random Effects: ([The whole goddamn model]|Participant, [The whole goddamn model]|Study)

# Set the most interesting contrast e.g. by defining synchronous or conversation as the baseline

# First we redefine the baseline.
phys_4$condition <- relevel(phys_4$condition, "Synchronous")

# Model change as a function of own and other previous state 

# This is an alternative syntax than just "Change ~ (HR_self + HR_other)*condition" but the output is easier to read, and the model is essentially the same.

HR_model <- lm(HR_change_self ~ 0 + condition + (HR_self + HR_other)*condition, data = phys_4)
HR_model2 <- lm(HR_change_self ~ 0 + condition + (HR_self + HR_other):condition, data = phys_4)

summary(HR_model)
object <- summary(HR_model2)

# [Insert interpretation here]

# Within each condition, all values for HR_self are negative, as expected. The higher your HR_self is, the more it causes the heart rate change to slow decrease / slow down.


### Adding random effects (Buckle in! It's going to be a bumpy ride!)
HR_model3 <- lmerTest::lmer(HR_change_self ~ 0 + condition +(HR_self + HR_other)*condition + (0+condition|participant), data = phys_4, control = lmerControl(optimizer = "nloptwrap", calc.derivs = FALSE,  optCtrl = list(ftol_abs = 1e-10, xtol_abs = 1e-10, maxeval=10000)))

summary(HR_model3)


anova(HR_model3, HR_model) # This didn't significantly improve the model



### See further down for this model with surrogate pairs.

```


## Now we need to create control baselines.

First shuffled controls, then surrogate pairs.

### Creating controls: shuffled controls

Shuffled controls break the temporal dependencies of time-series by shuffling the value within one time-series. This ensures the "coordination" observed is not due to the actual values in the series and not their sequence.
Tip: sample() is your friend, but make sure to shuffle things within participant/condition and not throughout the whole dataset
 
```{r}

# Create a shuffled dataset

set.seed(7)

shuffled <- phys_4

phys_4$Type <- "Real"

shuffled$HR_self = sample(shuffled$HR_self)
shuffled$HR_other = sample(shuffled$HR_other)
shuffled$HR_fut = sample(shuffled$HR_fut)
shuffled$HR_change_self = shuffled$HR_fut - shuffled$HR_self

# Concatenate it to the original dataset (and remember to have a column telling you which is which)

phys_4$Type <- "Real"
shuffled$Type <- "Fake"

p4_shuf <- rbind(phys_4, shuffled)

# Create the same models as in the previous chunk, but adding an interaction by shuffled vs. real

HR_model2 <- lm(HR_change_self ~ (HR_self + HR_other)*condition, data = p4_shuf)
summary(HR_model2)

Resp_model2 <- lm(Resp_change_self ~ (Resp_self + Resp_other)*condition, data = p4_shuf)

summary(Resp_model2)


```
 
 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair)

```{r}
pacman::p_load(pacman,tidyverse, stringr, stringi, groupdata2)

# Identify unique pairs within a given study (to keep things manageable) and create list of possible surrogate pairs (e.g. individual 1 from pair 1 and individual 2 from pair 2)


good_data <- read.csv("Bad Data Removed.csv")

good_data <- good_data %>% 
  group_by(group, condition) %>% 
  mutate(HR1fut = lead(HR1,1), #lead does the same as lag
         HR2fut = lead(HR2,1), 
         Resp1fut = lead(Resp1,1), 
         Resp2fut = lead(Resp2,1),
         HR1change = HR1fut - HR1,
         HR2change = HR2fut - HR2,
         Resp1change = Resp1fut - Resp1,
         Resp2change = Resp2fut - Resp2)

d_surrogate <- good_data[0,]
Groups <-as.numeric(as.character(unique(good_data$group[good_data$study==4]))) # List all pairs
SurrogateList<-expand.grid(a = Groups, b = Groups) # Identifyall possiblecombinationsof 2 pairs
SurrogateList= subset(SurrogateList, a != b) # excludecombinationswith identicalpairs

for (i in 1:nrow(SurrogateList)){ # loop throughall combinations
  print(SurrogateList$a[i])
  x <-subset(good_data, group==SurrogateList$a[i]) # subset data from the firstpair
  y <-subset(good_data, group!=SurrogateList$a[i]) # subset data from the secondpair 
  group<-c(800 + ((1:4)*i)) # createnew pair id
  for (co in c("Synchronous","TurnTaking", "Conversation", "MovementCoop", "MovementGuided")){ # loop through conditions
    if (co%in% unique(x$condition) & co%in% unique(y$condition)){ # check that both pairs have the data for that condition
      z1 <-subset(x, condition==co) # subset only that condtion from first pair
      z2 <-subset(y, condition==co) # subset only that condtion from second pair
      if (nrow(z1) > nrow(z2)) {z1<-z1[1:nrow(z2),]}
      if (nrow(z2) > nrow(z1)) { z2<-z2[1:nrow(z1),]}
      w1 <-z1 %>% mutate(# assemblenew pair combiningthe 2 pairs
          HR2 = z2$HR2,Resp2 = z2$Resp2,HR2fut = z2$HR2fut, Resp2fut = z2$Resp2fut, HR2change = z2$HR2change, Resp2change = z2$Resp2change)
      if (nrow(d_surrogate) == 0) {
        d_surrogate <- w1}
      else {
        d_surrogate <- rbind(d_surrogate,w1)
      }
      
      } }}



# Make it into long format


# creating long data for HR change
d_HR_change_self <- gather(d_surrogate, participant, HR_change_self, # new vars
HR1change, HR2change) %>% #old vars
  select( # drop irrelevant vars
  time, HR_change_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#creating long data for Respiration
d_Resp_change_self <- gather(d_surrogate, participant, Resp_change_self, # new vars
Resp1change, Resp2change) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_change_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#heart rate self
d_HR_self <- gather(d_surrogate, participant, HR_self, # new vars
HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
  time, HR_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))


#future heart rate
d_HR_fut <- gather(d_surrogate, participant, HR_fut, # new vars
HR1fut, HR2fut) %>% #old vars
  select( # drop irrelevant vars
  time, HR_fut, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))



#Respiration self
d_Resp_self <- gather(d_surrogate, participant, Resp_self, # new vars
Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_self, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#future Respiration rate
d_Resp_fut <- gather(d_surrogate, participant, Resp_fut, # new vars
Resp1fut, Resp2fut) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_fut, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))


# Making a new dataframe in which HR1/2 and Resp1/2 are switched, to generate the "other" variables
d_surrogate_2 = d_surrogate
d_surrogate_2$HR1 = d_surrogate$HR2
d_surrogate_2$HR2 = d_surrogate$HR1
d_surrogate_2$Resp1 = d_surrogate$Resp2
d_surrogate_2$Resp2 = d_surrogate$Resp1

#HR other
d_HR_other <- gather(d_surrogate_2, participant, HR_other, # new vars
HR1, HR2) %>% #old vars
  select( # drop irrelevant vars
  time, HR_other, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))

#Resp other
d_Resp_other <- gather(d_surrogate_2, participant, Resp_other, # new vars
Resp1, Resp2) %>% #old vars
  select( # drop irrelevant vars
  time, Resp_other, participant, study, group, condition, trial) %>% mutate( # create unique participant ID
  participant = parse_number(as.character(group)) * 10 + parse_number(participant))


#merging dataframes
surrogate_all = cbind(d_HR_change_self, d_HR_self, d_HR_fut, d_HR_other, d_Resp_change_self, d_Resp_self, d_Resp_fut, d_Resp_other)
surrogate_all_2 = surrogate_all %>% select(time, HR_change_self, HR_self, HR_fut, HR_other, participant, study, group, condition, trial, Resp_change_self, Resp_self, Resp_fut, Resp_other)

surrogate_long <- surrogate_all_2
surrogate_long$Type <- "Surrogate"

write.csv(surrogate_long, "Surrogate Data, Long.csv")
surrogate_long <- read.csv("Surrogate Data, Long.csv")
surrogate_long[,1] <- NULL

# Binding real and surrogate data
phys_long <- read.csv("Bad Data Removed, Long.csv")
phys_long[,1] <- NULL
phys_long$Type <- "Real"


really_long <- rbind(phys_long, surrogate_long)

write.csv(really_long, "Real + Surrogate Data, Long.csv")

```

```{r}
pacman::p_load(pacman,tidyverse, stringr, stringi, groupdata2)
# Create models as in chunks above, but adding an interaction with the Real vs. Surrogate variable (exclude shuffled ones for simplicity)

really_long <- read.csv("Real + Surrogate Data, Long.csv")
really_long[,1] <- NULL

## We will only analyze data from Study 4, because that is easier.
really_long_4 <- really_long %>% filter(study == 4)

## Fixed Effects

### HR_self = Stability ## Random effect: Well-trained people have different stability since HR variability is actually healthy.
### HR_other = coupling. ## Random effect: Some pairs might actually hate eachother.
### condition = condition ## Random effect: If you hate the other person, the condition doesn't matter.


## Syntax Shenanigans

### lm() implicitly adds a "1+" to the syntax, which means "Pick an intercept"
### 0+ tells lme4() to not make several intercepts. In this case, we should say "0+condition + (HR_s + HR_o):condition"
### Then take the above and do: (0+condition + (HR_s+HR_o):condition):Type

## Random Effects: ([The whole goddamn model]|Participant, [The whole goddamn model]|Study)

# Set the most interesting contrast e.g. by defining synchronous or conversation as the baseline

# First we redefine the baseline.
really_long_4$condition <- relevel(really_long_4$condition, "Synchronous")

# Model change as a function of own and other previous state 

# This is an alternative syntax than just "Change ~ (HR_self + HR_other)*condition" but the output is easier to read, and the model is essentially the same.

HR_model4 <- lm(HR_change_self ~ (0 + condition + (HR_self + HR_other)*condition):Type, data = really_long_4)
HR_model5 <- lm(HR_change_self ~ (0 + condition + (HR_self + HR_other):condition):Type, data = really_long_4)
HR_model6 <- lm(HR_change_self ~ (0 + condition + (HR_self + HR_other):condition)*Type, data = really_long_4)


summary(HR_model4)
summary(HR_model5)
summary(HR_model6)


table_1 <- summary(HR_model4)
table_2 <- summary(HR_model5)
table_3 <- summary(HR_model6)




write.csv(table_1[["coefficients"]], "Table 1.csv")

write.csv(table_2[["coefficients"]], "Table 2.csv")

write.csv(table_3[["coefficients"]], "Table 3.csv")

# [Insert interpretation here]

# Within each condition, all values for HR_self are negative, as expected. The higher your HR_self is, the more it causes the heart rate change to slow decrease / slow down.


### Adding random effects (Buckle in! It's going to be a bumpy ride!)


HR_model7 <- lmer(HR_change_self ~ ((HR_self + HR_other):condition):Type + (1+condition|participant), data = really_long_4)

summary(HR_model7)

# Is there significant coordination at all? - Stability and coupling

# Does coordination differ between conditions? 

# Is there significantly more coordination in the data than in data with surrogate controls?


anova(HR_model6, HR_model5)

### Code - Basically complete. Under the assumption that:
## - The "Change_other" variables need not be tested. - Resp need not be explicitly tested. - It's okay to simplify random effects.

```

### Effects of Respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them